Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-07-13 article-models-ttinternal.tex
pre-weakening homomorphisms

@@

@@

Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$. Weakening will be a
CFT-homomorphism
\begin{equation}\label{cftwk}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation}
satisfying judgmental equalities which express abstracted analogues of
the rules in \autoref{comp-ww,comp-ew,comp-0w}. The
definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.

We begin with some trivial remarks to familiarize ourselves with the situation.
The morphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
$(\cftwkc{\cftalg{A}},\cftwkf{\cftalg{A}},\cftwkt{\cftalg{A}})$ consisting of
the morphisms displayed in the following diagram:
\begin{equation*}
\begin{tikzcd}
\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
  \ar[fib]{d}
  \ar{r}{\cftwkt{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
  \ar[fib]{d}
  \\
\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \ar[fib]{d}
  \ar{r}{\cftwkf{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \ar[fib]{d}
  \\
\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}[swap]{\cftwkc{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
\end{tikzcd}
\end{equation*}
In particular, we see that $\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to $\cftfamalg{\cftalg{A}}$ over
$\cftctxext$ in context $\Gamma$.

@@

>>>>

\begin{thm}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then the triple
\begin{equation*}
\cftidhom{\extalg{A}}\defeq\unfold{\cftidhom{\extalg{A}}}
\end{equation*}
is an extension homomorphism from $\cftalg{A}$ to $\cftalg{A}$ in context
$\Gamma$.
\end{thm}

@@

\begin{lem}
For any CFT-algebra $\cftalg{A}$, the extension homomorphism $\cftidhom{\cftalg{A}}$
is a CFT-homomorphism.
\end{lem}

@@

\begin{defn}
A pre-weakening algebra $\cftalg{A}$ in context $\Gamma$ is a CFT-algebra
$\cftalg{A}$ in context $\Gamma$ together with a CFT-homomorphism
\begin{equation}\label{cftwk}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation}
called the \emph{weakening of $\cftalg{A}$}.
\end{defn}

Weakening algebras will be pre-weakening algebras for which the weakening
satisfies judgmental equalities which express abstracted analogues of
the rules in \autoref{comp-ww,comp-ew,comp-0w}. The
definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.

We begin with some trivial remarks to familiarize ourselves with the situation
of pre-weakening algebras.
The CFT-homomorphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
$(\cftwkc{\cftalg{A}},\cftwkf{\cftalg{A}},\cftwkt{\cftalg{A}})$ consisting of
the morphisms displayed in the following diagram:
\begin{equation*}
\begin{tikzcd}
\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
  \ar[fib]{d}
  \ar{r}{\cftwkt{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
  \ar[fib]{d}
  \\
\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \ar[fib]{d}
  \ar{r}{\cftwkf{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \ar[fib]{d}
  \\
\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}[swap]{\cftwkc{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
\end{tikzcd}
\end{equation*}
In particular, we see that $\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to $\cftfamalg{\cftalg{A}}$ over
$\cftctxext$ in context $\Gamma$.

@@

\begin{defn}
Suppose $\cftalg{A}$ is a pre-weakening algebra in context $\Gamma$. Then we
define the pre-weakening algebra $\cftfamalg{\cftalg{A}}$ in context 
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
\begin{equation*}
\cftwk{\cftfamalg{\cftalg{A}}}
  \defeq \jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{\Gamma}{B}$ and
let $\jfam{\Gamma}{A}$. Then we define the pre-weakening algebra 
$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
taking
\begin{equation*}
\cftwk{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftwk{\cftalg{Q}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-weakening algebra 
$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
\begin{equation*}
\cftwk{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftwk{\cftalg{Q}}}.
\end{equation*}
\end{defn}

\subsubsection{Pre-weakening homomorphisms}
It is useful to start with the definition of pre-weakening homomorphisms. Those
are CFT-homomorphisms which also preserve weakening. Before we give the
definition, recall from \autoref{lem:jvcomp-cprojfstf} that we have a
commutative square
\begin{equation*}
\begin{tikzcd}
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  \ar{d}[swap]{\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  &
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{B}}}
  \ar{d}{\cprojfstf{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
  &
\cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation*}
of morphisms in context $\Gamma$ for every CFT-homomorphism $\cfthom{f}$ from
$\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$. Because weakening is the
same thing as pulling back along projection, we therefore see that the
judgmental equality
\begin{equation*}
\jfameq
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  { \ctxwk
      {\cftalgf{\cftalg{A}}}
      {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}}
    }
  { \jcomp{}
      { \cfthomf{\cfthom{f}}}
      { \jcomp{}
        {\cfthomc{\cfthom{f}}}
        {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
        }
    }
\end{equation*}
is valid for every CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$. This judgmental equality helps us seeing that the condition
on pre-weakening homomorphisms is well-typed.

\begin{defn}
Let $\cftalg{A}$ and $\cftalg{B}$ be pre-weakening algebras in context $\Gamma$.
A \emph{pre-weakening homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context
$\Gamma$} is a CFT-homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$ for which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamhom{\cfthom{f}}}}
  \ar{d}[swap]{\cftwk{\cftalg{A}}}
  &
\jcomp{}
  { \cfthomf{\cfthom{f}}}
  { \jcomp{}
    {\cfthomc{\cfthom{f}}}
    {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
    }
  \ar{d}
    { \jcomp{}
        {\cfthomf{\cfthom{f}}}
        {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftwk{\cftalg{B}}}}
      }
  \\
\cftfamalg{\cftfamalg{\cftalg{A}}}
  \ar{r}[swap]{\cftfamhom{\cftfamhom{\cfthom{f}}}}
  &
\jcomp{}
  {\cfthomf{\cfthom{f}}}
  {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}
\end{tikzcd}
\end{equation*}
of CFT-algebras in context 
$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ commutes
judgmentally.
\end{defn}

\begin{rmk}
In this remark we explain what it means that weakening is a pre-weakening
homomorphism. Let $\cftalg{A}$ be a pre-weakening algebra. Then its weakening
$\cftwk{\cftalg{A}}$ is a CFT-homomorphism from
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
$\cftfamalg{\cftfamalg{\cftalg{A}}}$. Therefore, $\cftwk{\cftalg{A}}$ is a
pre-weakening homomorphism if the diagram
\begin{equation*}
\begin{tikzcd}[column sep=8em]
\ctxwk
  {\cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  {\cftfamalg{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  \ar{r}
    { \ctxwk
        {\cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \ar{d}[swap]{\cftwk{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  &
\jcomp{}
  { \cftwkf{\cftalg{A}}}
  { \jcomp{}
    {\cftwkc{\cftalg{A}}}
    {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
    }
  \ar{d}
    { \jcomp{}
        { \cftwkf{\cftalg{A}}}
        { \jcomp{}
            { \cftwkc{\cftalg{A}}}
            { \cftwk{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
      }
  \\
\cftfamalg{\cftfamalg{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  \ar{r}[swap]
    { \cftfamhom
        { \cftfamhom
            { \cftwk{\cftalg{A}}}
          }
      }
  &
\jcomp{}
  { \cftwkf{\cftalg{A}}}
  { \jcomp{}
      {\cftwkc{\cftalg{A}}}
      {\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
    }
\end{tikzcd}
\end{equation*}
Both the domain and the codomain are
again pre-weakening algebras: the weakening of
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ is
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamhom{\cftwk{\cftalg{A}}}}$; the weakening of
$\cftfamalg{\cftfamalg{\cftalg{A}}}$ is
$\cftfamhom{\cftfamhom{\cftwk{\cftalg{A}}}}$. Therefore, the condition that
the weakening of $\cftalg{A}$ is a pre-weakening homomorphism becomes that
the diagram
\end{rmk}
